--
title: "PAN Physical Activity and Cortical Microstructure Analysis"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
    code_folding: hide
params:
  data_file: "/Users/danielcallow/Downloads/data_export_2026-01-26 16_08_34.090270.csv"
  output_dir: "PAN_outputs"
---

This report focuses on the analyses reported in the PAN manuscript, including
Tables 1 and 2, Figure 1, secondary analyses, and sensitivity analyses. Set `data_file` and `output_dir` in the parameters above before knitting. Existing primary filtering, recoding, standardization, and model specifications
are preserved. Separate-predictor, volume-adjustment, and MoCA-exclusion models
are added from the manuscript Methods; their numerical agreement requires verification.

The final complete-case sample is `df_pr`. Outcomes are standardized before
filtering, and predictors after eligibility filtering but before complete-case
selection, matching the original script. Figure confidence bands use conventional
OLS inference; the main result tables use HC3 standard errors.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = TRUE)
required_packages <- c(
  "tidyverse", "janitor", "broom", "car", "gtsummary",
  "gt", "lmtest", "sandwich", "patchwork"
)
missing_packages <- required_packages[
  !vapply(required_packages, requireNamespace, logical(1), quietly = TRUE)
]
if (length(missing_packages)) {
  stop("Install these packages before knitting: ",
       paste(missing_packages, collapse = ", "), call. = FALSE)
}
invisible(lapply(required_packages, library, character.only = TRUE))
Sys.setenv(VROOM_CONNECTION_SIZE = 8388608)
output_dir <- params$output_dir
dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)
```


# Import data

```{r import-data}
file_path <- params$data_file
if (!file.exists(file_path)) stop("Data file not found: ", file_path, call. = FALSE)

df_raw <- read_csv(
  file_path,
  skip = 9,
  show_col_types = FALSE,
  guess_max = 50000,
  progress = FALSE
) %>%
  clean_names()

cat("Rows:", nrow(df_raw), "\n")
cat("Columns:", ncol(df_raw), "\n")
names(df_raw)
glimpse(df_raw)
```


# Vascular risk and ethnicity recodes

Missing vascular risk components are coded as zero in the original analysis; this behavior is preserved. The derived `hispanic_latino_bin` is retained, while the models use `hispanic_latino_hml`.

```{r vascular-risk}
df_raw <- df_raw %>%
  mutate(
    bmi = (bio_weight / (bio_height^2)) * 703
  )

df_raw <- df_raw %>%
  mutate(
    vrf_obesity = if_else(bmi > 30, 1L, 0L, missing = 0L),

    vrf_diabetes = if_else(
      health_medical_type_1_diabetes == 1 |
      health_medical_type_2_diabetes == 1,
      1L, 0L, missing = 0L
    ),

    vrf_smoking = if_else(
      health_medical_smoking == 1,
      1L, 0L, missing = 0L
    ),

    vrf_hypertension = if_else(
      health_medical_hypertension == 1,
      1L, 0L, missing = 0L
    ),

    vrf_cholesterol = if_else(
      cholesterol > 200,
      1L, 0L, missing = 0L
    )
  )

df_raw <- df_raw %>%
  mutate(
    vascular_risk_count =
      vrf_obesity +
      vrf_diabetes +
      vrf_smoking +
      vrf_hypertension +
      vrf_cholesterol,

    vascular_risk_cat = case_when(
      vascular_risk_count == 0 ~ "0",
      vascular_risk_count == 1 ~ "1",
      vascular_risk_count >= 2 ~ "2+",
      TRUE ~ NA_character_
    ),
    vascular_risk_cat = factor(
      vascular_risk_cat,
      levels = c("0", "1", "2+")
    )
  )

df_raw <- df_raw %>%
  mutate(
    hispanic_latino_bin = if_else(
      hispanic_latino == "Yes",
      1L,
      0L
    ),
    hispanic_latino_bin = factor(
      hispanic_latino_bin,
      levels = c(0, 1),
      labels = c("No / Not Hispanic", "Hispanic")
    )
  )
```


# Imaging and demographic variables

Education and race remain factors, matching the source code. Education is therefore modeled categorically, despite the manuscript describing an ordinal coding.

```{r derived-variables}
df_raw <- df_raw %>%
  mutate(
    site = factor(site),

    # Normalized cortical volume: cortical gray matter volume / eTIV
    cortexvol_norm_etiv =
      coree_chen_anat_measures_freesurfer_aseg_cortexvol /
      coree_chen_anat_measures_freesurfer_aseg_etiv,

    extra_neurite_fraction_cortex =
      pmax(
        1 -
          coree_chen_dwi_measures_noddi_gm_gmcort_fit_ndi_cerebral_cortex -
          coree_chen_dwi_measures_noddi_gm_gmcort_fit_fwf_cerebral_cortex,
        0
      ),

    edu_cat = case_when(
      highest_education_level_completed %in% c(
        "Some High School",
        "High School Diploma (Baccalaureate)"
      ) ~ 0,
      highest_education_level_completed == "Some College (Some University)" ~ 1,
      highest_education_level_completed == "College Degree (University Graduate)" ~ 2,
      highest_education_level_completed ==
        "Post Graduate Degree (Masters or Doctorate)" ~ 3,
      TRUE ~ NA_real_
    ),

    race_cat = case_when(
      race == "Black or African American" ~ 1,
      race == "White" ~ 2,
      TRUE ~ 0
    )
  ) %>%
  mutate(
    edu_cat = factor(
      edu_cat,
      levels = 0:3,
      labels = c(
        "HS or less",
        "Some college",
        "College degree",
        "Post-graduate"
      )
    ),
    race_cat = factor(
      race_cat,
      levels = c(0, 1, 2),
      labels = c(
        "Other / Not White or Black",
        "Black",
        "White"
      )
    )
  )
```


# Standardize outcomes

```{r standardize-outcomes}
by_site_outcomes <- c(
  "coree_chen_dwi_measures_noddi_gm_gmcort_fit_fwf_cerebral_cortex",
  "coree_chen_dwi_measures_noddi_gm_gmcort_fit_ndi_cerebral_cortex",
  "coree_chen_dwi_measures_noddi_gm_gmcort_fit_odi_cerebral_cortex",
  "coree_chen_dwi_measures_tensor_gm_gmcort_desc_adc_cerebral_cortex",
  "cortexvol_norm_etiv"
)

df_raw <- df_raw %>%
  group_by(site) %>%
  mutate(
    across(
      all_of(by_site_outcomes),
      ~ as.numeric(scale(.x)),
      .names = "{.col}_z_site"
    )
  ) %>%
  ungroup()

outcomes_y <- paste0(by_site_outcomes, "_z_site")

# 1B) Create globally z-scored outcomes for sensitivity analyses
#     These are NOT standardized within site

df_raw <- df_raw %>%
  mutate(
    across(
      all_of(by_site_outcomes),
      ~ as.numeric(scale(.x)),
      .names = "{.col}_z_global"
    )
  )

outcomes_global_z <- paste0(by_site_outcomes, "_z_global")
```


# Analysis specification and eligibility

```{r analysis-specification}
predictors <- c(
  "mvpa_min",
  "mean_daily_10hr_steps"
)

covariates <- c(
  "age_hml",
  "sex_hml",
  "edu_cat",
  "race_cat",
  "apoe_e4_carrier",
  "days_of_recording",
  "site",
  "hispanic_latino_hml",
  "vascular_risk_cat"
)

# 3) Build analysis dataset
df_analysis <- df_raw %>%
  filter(
    days_of_recording >= 6,
    days_of_recording <= 31,
    daily_monitoring_hr >= 20
  ) %>%
  mutate(
    sex_hml = factor(sex_hml),
    site = factor(site),

    apoe_e4_carrier = case_when(
      coreg_huentelman_apoe4_apoe_status %in% c("e3e4", "e4e4", "e2e4") ~ 1,
      coreg_huentelman_apoe4_apoe_status %in% c("e2e2", "e2e3", "e3e3") ~ 0,
      TRUE ~ NA_real_
    ),

    mvpa_min = mean_moderate_intensity_duration_minute +
      mean_vigorous_intensity_duration_minute,

    cortexvol_norm_etiv =
      coree_chen_anat_measures_freesurfer_aseg_cortexvol /
      coree_chen_anat_measures_freesurfer_aseg_etiv
  ) %>%
  filter(
    mean_daily_10hr_steps > 100,
    mean_daily_10hr_steps <= 25000,
    mvpa_min >= 0,
    #moca_total > 23,
    coree_chen_dwi_measures_tensor_gm_gmcort_desc_adc_cerebral_cortex < 0.0016,
    is.finite(cortexvol_norm_etiv),
    coree_chen_anat_measures_freesurfer_aseg_etiv > 0,
    coree_chen_anat_measures_freesurfer_aseg_cortexvol > 0
  ) %>%
  mutate(
    # skew-robust sensitivity versions
    mvpa_min_log = log1p(mvpa_min),
    mean_daily_10hr_steps_log = log1p(mean_daily_10hr_steps)
  )

# 4) Standardize predictors globally
df_analysis <- df_analysis %>%
  mutate(
    across(
      all_of(predictors),
      ~ as.numeric(scale(.x)),
      .names = "{.col}_z"
    ),
    across(
      c(mvpa_min_log, mean_daily_10hr_steps_log),
      ~ as.numeric(scale(.x)),
      .names = "{.col}_z"
    )
  )

predictors_z <- c("mvpa_min_z", "mean_daily_10hr_steps_z")
predictors_log_z <- c("mvpa_min_log_z", "mean_daily_10hr_steps_log_z")
outcomes_z <- outcomes_y
```


# Model helper

For multilevel factors, VIF output retains GVIF, degrees of freedom, and adjusted GVIF. Sorting uses adjusted GVIF instead of treating raw GVIF as ordinary VIF; the original 5/10 screening flags are retained and should be interpreted with the metric shown.

```{r model-helper}
run_lm <- function(outcome, predictors_z, covariates, data, robust_type = "HC3") {
  f <- as.formula(
    paste(outcome, "~", paste(c(predictors_z, covariates), collapse = " + "))
  )

  lm_fit <- lm(f, data = data)

  # conventional OLS output
  out_tbl_ols <- broom::tidy(lm_fit, conf.int = TRUE) %>%
    mutate(
      outcome = outcome,
      se_type = "OLS"
    )

  # robust HC3 output
  vcov_rob <- sandwich::vcovHC(lm_fit, type = robust_type)
  ct_rob   <- lmtest::coeftest(lm_fit, vcov. = vcov_rob)

  out_tbl_rob <- broom::tidy(ct_rob) %>%
    mutate(
      conf.low = estimate - qt(0.975, df = df.residual(lm_fit)) * std.error,
      conf.high = estimate + qt(0.975, df = df.residual(lm_fit)) * std.error,
      outcome = outcome,
      se_type = robust_type
    ) %>%
    dplyr::select(term, estimate, std.error, statistic, p.value, conf.low, conf.high, outcome, se_type)

  vif_values <- car::vif(lm_fit)
  if (is.matrix(vif_values)) {
    vif_tbl <- as.data.frame(vif_values) %>%
      tibble::rownames_to_column("term") %>%
      mutate(vif = .data[["GVIF^(1/(2*Df))"]], vif_metric = "Adjusted GVIF")
  } else {
    vif_tbl <- tibble(term = names(vif_values), vif = as.numeric(vif_values),
                      vif_metric = "VIF")
  }
  vif_tbl <- vif_tbl %>% mutate(outcome = outcome)

  list(
    model = lm_fit,
    tidy_ols = out_tbl_ols,
    tidy_robust = out_tbl_rob,
    vif = vif_tbl
  )
}
```


# Final complete-case sample

```{r complete-cases}
analysis_vars <- unique(c(
  outcomes_z,          # within-site z outcomes
  outcomes_global_z,   # global z outcomes for sensitivity analysis
  by_site_outcomes,
  predictors,
  predictors_z,
  predictors_log_z,
  covariates,
  "daily_monitoring_hr",
  "moca_total",
  "cortexvol_norm_etiv",
  "mvpa_min_log",
  "mean_daily_10hr_steps_log"
))

df_pr <- df_analysis %>%
  dplyr::select(all_of(analysis_vars)) %>%
  tidyr::drop_na() %>%
  dplyr::filter(if_all(where(is.numeric), is.finite))

cat("N in df_analysis:", nrow(df_analysis), "\n")
cat("N in df_pr:", nrow(df_pr), "\n")
```


# Bivariate associations

```{r bivariate-associations}
df_biv <- df_pr %>%
  dplyr::select(
    mvpa_min,
    mean_daily_10hr_steps,
    coree_chen_dwi_measures_tensor_gm_gmcort_desc_adc_cerebral_cortex,
    coree_chen_dwi_measures_noddi_gm_gmcort_fit_fwf_cerebral_cortex,
    coree_chen_dwi_measures_noddi_gm_gmcort_fit_ndi_cerebral_cortex,
    coree_chen_dwi_measures_noddi_gm_gmcort_fit_odi_cerebral_cortex
    #noddi_wm_wmintra_desc_icvf_cerebral_white_matter,
    #noddi_wm_wmintra_desc_isovf_cerebral_white_matter,
    #noddi_wm_wmintra_desc_od_cerebral_white_matter,
    #tensor_wm_wmintra_desc_adc_cerebral_white_matter,
    #tensor_wm_wmintra_desc_fa_cerebral_white_matter
  ) %>%
  rename(
    MVPA = mvpa_min,
    Steps = mean_daily_10hr_steps,
    MD_GM = coree_chen_dwi_measures_tensor_gm_gmcort_desc_adc_cerebral_cortex,
    FWF_GM = coree_chen_dwi_measures_noddi_gm_gmcort_fit_fwf_cerebral_cortex,
    NDI_GM = coree_chen_dwi_measures_noddi_gm_gmcort_fit_ndi_cerebral_cortex,
    ODI_GM = coree_chen_dwi_measures_noddi_gm_gmcort_fit_odi_cerebral_cortex
    #NDI_WM = noddi_wm_wmintra_desc_icvf_cerebral_white_matter,
    #FWF_WM = noddi_wm_wmintra_desc_isovf_cerebral_white_matter,
    #ODI_WM = noddi_wm_wmintra_desc_od_cerebral_white_matter,
    #MD_WM = tensor_wm_wmintra_desc_adc_cerebral_white_matter,
    #FA_WM = tensor_wm_wmintra_desc_fa_cerebral_white_matter
  )

vars_biv <- names(df_biv)

get_cor_results <- function(data, vars, method = "pearson") {
  combn(vars, 2, simplify = FALSE) %>%
    purrr::map_dfr(function(v) {
      tt <- suppressWarnings(
        cor.test(
          data[[v[1]]],
          data[[v[2]]],
          method = method,
          exact = FALSE
        )
      )

      out <- broom::tidy(tt) %>%
        mutate(
          var1 = v[1],
          var2 = v[2],
          comparison = paste(v[1], "vs", v[2]),
          method = method,
          .before = 1
        )

      # make sure CI columns exist even when cor.test/broom does not return them
      if (!"conf.low" %in% names(out)) out$conf.low <- NA_real_
      if (!"conf.high" %in% names(out)) out$conf.high <- NA_real_

      out
    }) %>%
    mutate(
      p_fdr = p.adjust(p.value, method = "fdr")
    ) %>%
    dplyr::select(
      method, comparison, var1, var2,
      estimate, conf.low, conf.high, p.value, p_fdr
    )
}

biv_pearson <- get_cor_results(df_biv, vars_biv, method = "pearson") %>%
  mutate(
    across(c(estimate, conf.low, conf.high), ~ round(.x, 3)),
    p.value = signif(p.value, 3),
    p_fdr = signif(p_fdr, 3)
  )

biv_pearson_reported <- biv_pearson %>%
  filter(
    (var1 == "MVPA" & var2 == "Steps") |
      (var1 %in% c("MD_GM", "FWF_GM", "NDI_GM", "ODI_GM") &
         var2 %in% c("MD_GM", "FWF_GM", "NDI_GM", "ODI_GM"))
  )
biv_pearson_reported

```


# Participant characteristics

```{r participant-characteristics}
df_table1 <- df_pr %>%
  mutate(
    sex_hml = case_when(
      as.character(sex_hml) %in% c("0", "0.0", "Female", "female", "F", "f") ~ "Female",
      as.character(sex_hml) %in% c("1", "1.0", "Male", "male", "M", "m") ~ "Male",
      TRUE ~ as.character(sex_hml)
    ),
    sex_hml = factor(sex_hml),

    apoe_e4_carrier = factor(
      apoe_e4_carrier,
      levels = c(0, 1),
      labels = c("Non-carrier", "Carrier")
    ),

    hispanic_latino_hml = case_when(
      as.character(hispanic_latino_hml) %in% c("0", "0.0", "No", "No / Not Hispanic", "Not Hispanic") ~ "No / Not Hispanic",
      as.character(hispanic_latino_hml) %in% c("1", "1.0", "Yes", "Hispanic") ~ "Hispanic",
      TRUE ~ as.character(hispanic_latino_hml)
    ),
    hispanic_latino_hml = factor(hispanic_latino_hml)
  )

table1 <- df_table1 %>%
  dplyr::select(
    age_hml,
    sex_hml,
    edu_cat,
    race_cat,
    hispanic_latino_hml,
    apoe_e4_carrier,
    vascular_risk_cat,
    days_of_recording,
    daily_monitoring_hr,
    mvpa_min,
    mean_daily_10hr_steps,
    cortexvol_norm_etiv,
    moca_total
  ) %>%
  tbl_summary(
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    digits = list(
      all_continuous() ~ 2,
      mvpa_min ~ 1,
      mean_daily_10hr_steps ~ 0
    ),
    label = list(
      age_hml ~ "Age, years",
      sex_hml ~ "Sex",
      edu_cat ~ "Education",
      race_cat ~ "Race",
      hispanic_latino_hml ~ "Ethnicity",
      apoe_e4_carrier ~ "APOE ε4 carrier",
      vascular_risk_cat ~ "Vascular risk category",
      days_of_recording ~ "Valid recording days",
      daily_monitoring_hr ~ "Monitoring hours/day",
      mvpa_min ~ "MVPA, min/day",
      mean_daily_10hr_steps ~ "Mean daily step count",
      moca_total ~ "MoCA score"
    ),
    missing = "no"
  ) %>%
  bold_labels() %>%
  modify_caption(
    paste0("**Table 1. Participant characteristics (N = ", nrow(df_pr), ")**")
  ) %>%
  as_gt()

table1
```


# Primary regression results

The original FDR adjustment spans both activity predictors across all five outcomes and is retained as `p_fdr`. NODDI-only adjustments are also displayed below by activity predictor; the manuscript does not specify whether both predictors belong to one family. MD is not a NODDI-derived measure, although Table 2 marks it with the NODDI FDR footnote.

```{r primary-models}
fits <- purrr::map(
  outcomes_z,
  ~ run_lm(
    outcome      = .x,
    predictors_z = predictors_z,
    covariates   = covariates,
    data         = df_pr,
    robust_type  = "HC3"
  )
)
names(fits) <- outcomes_z

# Collect results ---------------------------------------------------------
all_tidy_ols    <- purrr::map_dfr(fits, "tidy_ols")
all_tidy_robust <- purrr::map_dfr(fits, "tidy_robust")
all_vif         <- purrr::map_dfr(fits, "vif")

# Primary manuscript results: HC3 robust SEs -----------------------------
results_main <- all_tidy_robust %>%
  filter(term %in% predictors_z) %>%
  dplyr::select(outcome, term, estimate, std.error, conf.low, conf.high, p.value, se_type) %>%
  mutate(
    p_fdr = p.adjust(p.value, method = "fdr")
  ) %>%
  arrange(outcome, p.value)

results_main

# Explicit NODDI-only families, separately for each activity predictor.
results_noddi_fdr <- all_tidy_robust %>%
  filter(term %in% predictors_z, grepl("noddi", outcome)) %>%
  group_by(term) %>%
  mutate(p_fdr_noddi = p.adjust(p.value, method = "BH")) %>%
  ungroup()
results_noddi_fdr

# Full coefficients support the covariate findings reported in the Results.
results_covariates <- all_tidy_robust %>%
  filter(!term %in% c("(Intercept)", predictors_z))
results_covariates

```


# Outcome names
outcome_md <- "coree_chen_dwi_measures_tensor_gm_gmcort_desc_adc_cerebral_cortex_z_site"

outcome_fwf <- "coree_chen_dwi_measures_noddi_gm_gmcort_fit_fwf_cerebral_cortex_z_site"

# Reusable prediction-grid function
make_pred_grid <- function(data, mvpa_var = "mvpa_min_z", n = 150) {

  mvpa_seq <- seq(
    min(data[[mvpa_var]], na.rm = TRUE),
    max(data[[mvpa_var]], na.rm = TRUE),
    length.out = n
  )

  # Numeric covariates: mean
  typical <- data %>%
    summarise(
      across(
        everything(),
        ~ if (is.numeric(.x)) mean(.x, na.rm = TRUE) else NA
      )
    )

  nd <- typical[rep(1, n), , drop = FALSE]
  nd[[mvpa_var]] <- mvpa_seq

  # Factor covariates: modal/most common level
  factor_vars <- names(data)[sapply(data, is.factor)]

  for (v in factor_vars) {
    modal_level <- names(which.max(table(data[[v]])))
    nd[[v]] <- factor(modal_level, levels = levels(data[[v]]))
  }

  nd
}

# Reusable plot function
plot_mvpa_prediction <- function(
    outcome,
    model_list,
    data,
    y_label,
    plot_title = ""
) {

  m <- model_list[[outcome]]$model

  nd <- make_pred_grid(
    data = data,
    mvpa_var = "mvpa_min_z",
    n = 150
  )

  pred <- predict(
    m,
    newdata = nd,
    interval = "confidence",
    level = 0.95
  )

  pred_df <- bind_cols(
    nd %>% dplyr::select(mvpa_min_z),
    as_tibble(pred)
  )

  ggplot(pred_df, aes(x = mvpa_min_z, y = fit)) +
    geom_ribbon(
      aes(ymin = lwr, ymax = upr),
      alpha = 0.20,
      linewidth = 0
    ) +
    geom_line(linewidth = 1.2) +
    labs(
      x = "MVPA, standardized",
      y = y_label,
      title = plot_title
    ) +
    theme_classic(base_size = 12)
}

# MD plot
p_mvpa_md <- plot_mvpa_prediction(
  outcome = outcome_md,
  model_list = fits,
  data = df_pr,
  y_label = "Cortical MD, z-scored within site",
  plot_title = "MVPA and cortical mean diffusivity"
)

# FWF plot
p_mvpa_fwf <- plot_mvpa_prediction(
  outcome = outcome_fwf,
  model_list = fits,
  data = df_pr,
  y_label = "Cortical FWF, z-scored within site",
  plot_title = "MVPA and cortical free-water fraction"
)

print(p_mvpa_md)
print(p_mvpa_fwf)

ggsave(
  filename = file.path(output_dir, "mvpa_cortical_md_prediction.png"),
  plot = p_mvpa_md,
  width = 5.5,
  height = 4,
  dpi = 300
)

ggsave(
  filename = file.path(output_dir, "mvpa_cortical_fwf_prediction.png"),
  plot = p_mvpa_fwf,
  width = 5.5,
  height = 4,
  dpi = 300
)

p_mvpa_md + p_mvpa_fwf +
  plot_layout(ncol = 2)

p_mvpa_md_fwf <- p_mvpa_md + p_mvpa_fwf +
  plot_layout(ncol = 2)

ggsave(
  filename = file.path(output_dir, "mvpa_md_fwf_prediction_combined.png"),
  plot = p_mvpa_md_fwf,
  width = 10,
  height = 4,
  dpi = 300
)
```


# Partial regression figures

```{r partial-figures}
# Outcomes to plot
outcome_md <- "coree_chen_dwi_measures_tensor_gm_gmcort_desc_adc_cerebral_cortex_z_site"

outcome_fwf <- "coree_chen_dwi_measures_noddi_gm_gmcort_fit_fwf_cerebral_cortex_z_site"

# Covariates and other PA predictor to adjust for
rhs_no_mvpa <- c(
  setdiff(predictors_z, "mvpa_min_z"),
  covariates
)

# Function to create partial regression dataframe
make_partial_df <- function(data, outcome, predictor = "mvpa_min_z", covariates_adjust) {

  f_y <- as.formula(
    paste(outcome, "~", paste(covariates_adjust, collapse = " + "))
  )

  f_x <- as.formula(
    paste(predictor, "~", paste(covariates_adjust, collapse = " + "))
  )

  vars_needed <- unique(c(outcome, predictor, covariates_adjust))

  df_partial <- data %>%
    dplyr::select(all_of(vars_needed)) %>%
    tidyr::drop_na() %>%
    dplyr::filter(if_all(where(is.numeric), is.finite))

  ry <- resid(lm(f_y, data = df_partial))
  rx <- resid(lm(f_x, data = df_partial))

  tibble(
    mvpa_partial = rx,
    outcome_partial = ry
  )
}

# Function to plot partial regression in your desired visual style
plot_partial_mvpa <- function(partial_df, y_label) {

  ggplot(partial_df, aes(x = mvpa_partial, y = outcome_partial)) +
    geom_point(
      color = "black",
      alpha = 0.55,
      size = 2.2
    ) +
    geom_smooth(
      method = "lm",
      se = TRUE,
      color = "#2F55F5",
      fill = "gray70",
      linewidth = 1.3
    ) +
    labs(
      x = "Moderate to Vigorous Physical Activity",
      y = y_label
    ) +
    theme_classic(base_size = 14) +
    theme(
      panel.background = element_rect(),
      plot.background = element_rect(fill = "white", color = NA),
      axis.line = element_line(color = "black", linewidth = 0.8),
      axis.ticks = element_line(color = "black", linewidth = 0.8),
      axis.text = element_text(color = "black", size = 12),
      axis.title = element_text(color = "black", size = 16)
    )
}

# Figure 1: model-consistent residualization includes step count.
partial_md <- make_partial_df(df_pr, outcome_md, "mvpa_min_z", rhs_no_mvpa)
p_partial_md <- plot_partial_mvpa(partial_md, "Cortical Mean Diffusivity")
print(p_partial_md)
ggsave(file.path(output_dir, "Figure1_mvpa_cortical_md.png"),
       plot = p_partial_md, width = 7, height = 4.8, dpi = 300)
```

# Sex interaction analyses

Sex interaction outputs retain conventional OLS inference from the original script.

```{r secondary-models}
run_lm_sex_int <- function(outcome, covariates, data) {

  f <- as.formula(
    paste(
      outcome, "~",
      paste(
        c(
          "mvpa_min_z * sex_hml",
          "mean_daily_10hr_steps_z * sex_hml",
          setdiff(covariates, "sex_hml")
        ),
        collapse = " + "
      )
    )
  )

  lm(f, data = data)
}

fits_sex <- purrr::map(
  outcomes_z,
  ~ run_lm_sex_int(
    outcome    = .x,
    covariates = covariates,
    data       = df_pr
  )
)

names(fits_sex) <- outcomes_z

results_sex_int <- purrr::map_dfr(
  fits_sex,
  ~ broom::tidy(.x, conf.int = TRUE),
  .id = "outcome"
)

# inspect all model terms first
results_sex_int %>%
  dplyr::distinct(term) %>%
  dplyr::arrange(term)

# extract both interaction terms regardless of order
results_sex_int_interactions <- results_sex_int %>%
  dplyr::filter(
    (stringr::str_detect(term, "mvpa_min_z") &
     stringr::str_detect(term, "sex_hml")) |
    (stringr::str_detect(term, "mean_daily_10hr_steps_z") &
     stringr::str_detect(term, "sex_hml"))
  ) %>%
  dplyr::mutate(
    p_fdr = p.adjust(p.value, method = "fdr")
  ) %>%
  dplyr::arrange(outcome, term)

results_sex_int_interactions %>%
  dplyr::select(outcome, term, estimate, std.error, conf.low, conf.high, p.value, p_fdr)
```



# Separate activity predictors

These models use the primary complete-case sample and covariates, with one activity
predictor at a time. HC3 inference follows the manuscript's regression approach.

```{r separate-predictors}
results_separate <- purrr::map_dfr(predictors_z, function(activity) {
  purrr::map_dfr(outcomes_z, function(outcome) {
    run_lm(outcome, activity, covariates, df_pr)$tidy_robust %>%
      filter(term == activity)
  })
})
results_separate
```

# Cortical volume adjustment

The Methods specify diffusion outcomes as dependent variables with normalized
cortical volume as an additional covariate. This implements that specification;
the Results phrase describing both imaging measures as predictors needs clarification.

```{r volume-adjustment}
diffusion_outcomes <- setdiff(outcomes_z, "cortexvol_norm_etiv_z_site")
results_volume_adjusted <- purrr::map_dfr(diffusion_outcomes, function(outcome) {
  run_lm(outcome, predictors_z,
         c(covariates, "cortexvol_norm_etiv"), df_pr)$tidy_robust %>%
    filter(term %in% predictors_z)
})
results_volume_adjusted
```

# Exclude possible cognitive impairment

The manuscript excludes MoCA scores below 24. Existing standardizations are retained,
so this tests exclusion without changing the units of the primary coefficients.

```{r moca-sensitivity}
df_no_impairment <- df_pr %>% filter(moca_total >= 24)
cat("Excluded with MoCA < 24:", nrow(df_pr) - nrow(df_no_impairment), "\n")
cat("Sensitivity sample:", nrow(df_no_impairment), "\n")
results_no_impairment <- purrr::map_dfr(outcomes_z, function(outcome) {
  run_lm(outcome, predictors_z, covariates, df_no_impairment)$tidy_robust %>%
    filter(term %in% predictors_z)
})
results_no_impairment
```

# Global standardization sensitivity analyses

```{r site-sensitivity}
fits_global_z <- purrr::map(
  outcomes_global_z,
  ~ run_lm(
    outcome      = .x,
    predictors_z = predictors_z,
    covariates   = covariates,  # includes site
    data         = df_pr,
    robust_type  = "HC3"
  )
)

names(fits_global_z) <- outcomes_global_z

results_global_z <- purrr::map_dfr(fits_global_z, "tidy_robust") %>%
  filter(term %in% predictors_z) %>%
  mutate(
    p_fdr = p.adjust(p.value, method = "fdr"),
    analysis = "Global-z outcomes + site covariate"
  ) %>%
  dplyr::select(
    analysis, outcome, term, estimate, std.error,
    conf.low, conf.high, p.value, p_fdr, se_type
  )

results_global_z

# 9B) Sensitivity: globally z-scored outcomes with PA × site interactions

run_lm_site_int <- function(outcome, covariates, data, robust_type = "HC3") {

  # Remove site from covariates because it is already included in the interaction terms
  covars_no_site <- setdiff(covariates, "site")

  f <- as.formula(
    paste(
      outcome, "~",
      paste(
        c(
          "mvpa_min_z * site",
          "mean_daily_10hr_steps_z * site",
          covars_no_site
        ),
        collapse = " + "
      )
    )
  )

  lm_fit <- lm(f, data = data)

  vcov_rob <- sandwich::vcovHC(lm_fit, type = robust_type)
  ct_rob <- lmtest::coeftest(lm_fit, vcov. = vcov_rob)

  tidy_rob <- broom::tidy(ct_rob) %>%
    mutate(
      conf.low = estimate - qt(0.975, df = df.residual(lm_fit)) * std.error,
      conf.high = estimate + qt(0.975, df = df.residual(lm_fit)) * std.error,
      outcome = outcome,
      se_type = robust_type
    ) %>%
    dplyr::select(
      term, estimate, std.error, statistic, p.value,
      conf.low, conf.high, outcome, se_type
    )

  list(
    model = lm_fit,
    tidy_robust = tidy_rob
  )
}

fits_site_int_global_z <- purrr::map(
  outcomes_global_z,
  ~ run_lm_site_int(
    outcome    = .x,
    covariates = covariates,
    data       = df_pr,
    robust_type = "HC3"
  )
)

names(fits_site_int_global_z) <- outcomes_global_z

results_site_int_global_z <- purrr::map_dfr(
  fits_site_int_global_z,
  "tidy_robust"
) %>%
  filter(
    stringr::str_detect(term, "mvpa_min_z:site") |
      stringr::str_detect(term, "mean_daily_10hr_steps_z:site")
  ) %>%
  mutate(
    p_fdr = p.adjust(p.value, method = "fdr"),
    analysis = "Global-z outcomes + PA × site interactions"
  ) %>%
  arrange(outcome, term)

results_site_int_global_z %>%
  dplyr::select(
    analysis, outcome, term, estimate, std.error,
    conf.low, conf.high, p.value, p_fdr
  )
```


# Reproducibility

```{r session-info}
sessionInfo()
```
